---
title: "GAM to model the proportion of time spent at a clicking period through depth and by whale"
author: "Carolina S. Marques"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
csl: mee.csl
subtitle: 
bibliography: references.bib
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(knitr)
library(ggplot2)
library(lubridate)
library(dplyr) ##necessary to the function filter 
library(mgcv)
library(MASS)
library(nlme)
library(RVAideMemoire)
library(car)
library(hrbrthemes) 
library(dplyr)
library(tidyverse)
library(Hmisc)
theme_set(theme_bw())
options(encoding = 'UTF-8')
```

# Introduction 

In this document we model the probability of being in a clicking period as a function of depth and whale id.


# Reading data 


```{r}
#read data obtained in "01_data_processing.rmd":
load("processed_data.rda")
sampledata<-sampledata[sampledata$rsecs>=0&!is.na(sampledata$clicks),]

```


#  GAM models 

For computational efficiency, the data was sub-sampled considering a systematic sampling with random start, selecting only one out of every 10 samples in the original data. 

```{r}
originaldata$Clicking1<-ifelse(originaldata$click_status=="Silent",0,1)
                         
originaldata1=originaldata[seq(1,nrow(originaldata),by=10),]

```


## Model as a function of depth by whale


```{r}
gamwhales<-gam(Clicking1~s(Depth,by=as.factor(whale),k=4),data=originaldata1,family=binomial(link="logit"))
```

### Summary of the model 

```{r}  
summary(gamwhales) 
```

### AIC

```{r}
AIC(gamwhales)
```

### Predicting the influence of both varibles in the data


```{r}

originaldata11<-data.frame(Depth=originaldata1$Depth,whale=originaldata1$whale)

datat1<-data.frame(pclick=predict(gamwhales,newdata = originaldata11,type="response"),Depth=originaldata11$Depth,whale=originaldata11$whale)#get the prediction

confintervals<-data.frame(confintervals=predict(gamwhales,newdata=(originaldata11),type="response",se.fit=T))#get the confidence intervals

datat1$bigfit<-confintervals$confintervals.fit+confintervals$confintervals.se.fit
datat1$smallfit<-confintervals$confintervals.fit-confintervals$confintervals.se.fit

datat1$whale<-factor(datat1$whale,levels=c("Freya","Thora","Mára","Frida","Eistla","Balder","Jonas","Mutti"))

```

### Plot of the results from the model / Figure 4

```{r}
ggplot(data=datat1,aes(x=Depth,y=pclick,color=whale,fill=whale))+geom_line(linewidth=1.2)+ylab("Predicted probabilty of being in \na clicking period")+xlab("Depth(m)")+
  geom_ribbon(aes(ymin = smallfit,ymax=bigfit,x=Depth), alpha = .4)+
  theme_bw()+  scale_color_manual(values=c("black", "blue4","dodgerblue1","turquoise4","green4","darkorchid4","coral4","darkorange"))+  scale_fill_manual(values=c("black", "blue4","dodgerblue1","turquoise4","green4","darkorchid4","coral4","darkorange"))+labs(fill="",color="")+
  theme(
        axis.title.x = element_text(size = 20), 
        axis.text.x = element_text(size = 19),
        axis.text.y = element_text(size = 19),
        axis.title.y = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text= element_text(size=19),
        plot.title = element_text(size=20),
        strip.text.x = element_text(size = 21))
```

## Model as a function of depth


```{r}
gamwhales1<-gam(Clicking1~s(Depth,k=4),data=originaldata1,family=binomial(link="logit"))
```

### Summary of the model  

```{r}
summary(gamwhales1) 
```


### AIC

```{r}
AIC(gamwhales1)
```



### Predicting the influence of both variables in the data


```{r}

originaldata111<-data.frame(Depth=originaldata1$Depth)

datat11<-data.frame(pclick=predict(gamwhales1,newdata = originaldata111,type="response"),Depth=originaldata111$Depth)#get the prediction

confintervals1<-data.frame(confintervals=predict(gamwhales1,newdata=(originaldata111),type="response",se.fit=T))#get the confidence intervals

datat11$bigfit<-confintervals1$confintervals.fit+confintervals1$confintervals.se.fit
datat11$smallfit<-confintervals1$confintervals.fit-confintervals1$confintervals.se.fit

```

### Plot of the results from the model 

```{r}
ggplot(data=datat11,aes(x=Depth,y=pclick))+geom_line(linewidth=1.2)+ylab("Predicted probabilty of being in \na clicking period")+xlab("Depth(m)")+
  geom_ribbon(aes(ymin = smallfit,ymax=bigfit,x=Depth), alpha = .4)+
  theme_bw()+  scale_color_manual(values=c("black", "blue4","dodgerblue1","turquoise4","green4","darkorchid4","coral4","darkorange"))+  scale_fill_manual(values=c("black", "blue4","dodgerblue1","turquoise4","green4","darkorchid4","coral4","darkorange"))+labs(fill="",color="")+
  theme(
        axis.title.x = element_text(size = 20), 
        axis.text.x = element_text(size = 19),
        axis.text.y = element_text(size = 19),
        axis.title.y = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text= element_text(size=19),
        plot.title = element_text(size=20),
        strip.text.x = element_text(size = 21))
```


## Comparing both models

```{r}
data.frame("Response variable"=rep("Clicking/No clicking",2),"Explanatory variables"=c("depth by narwhal ID","depth"),"AIC"=c(round(AIC(gamwhales),3),round(AIC(gamwhales1),3)))
```

## Saving the best model and its results

```{r}
save(file="model_results_click_no_click.rda",list=c("gamwhales","datat1"))

```






